---
title: "Linear Models"
output: github_document
---

```{r setup}
library(tidyverse)
library(viridis)

library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE, 
  fig.width = 8, 
  fig.height = 6, 
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "virids"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Simulate a dataset 

* set.seed() : important way to make sure to get the same thing every time when doing a process that involves randomness 

```{r}
set.seed(1)

nonlin_df = 
  tibble(
    id = 1:100,
    x = runif(100, 0, 1), 
    y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
  )

nonlin_df %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point()
```

Cross validation - by hand

* create splits by hand; plot; fit some models 
* train_df taking 80% (80/100 observations)
* test_df needs to take the remaining 20%, not in the train_df -- looking for overlap by "id" variable 

```{r}
train_df = sample_n(nonlin_df, 80)
test_df = anti_join(nonlin_df, train_df, by = "id")

ggplot(train_df, aes(x = x, y = y)) + geom_point() + geom_point(data = test_df, color = "red")
```

fit models: 

```{r}
linear_mod = lm(y ~x, data = train_df)
smooth_mod = mgcv::gam(y ~ s(x), data = train_df)
wiggly_mod = mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df)
```

plot the results 

* linear model not complex enough
* wiggly model is more complex than smooth model
* putting too much into model 

```{r}
train_df %>% 
  add_predictions(linear_mod) %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred))

train_df %>% 
  add_predictions(smooth_mod) %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred))

train_df %>% 
  add_predictions(wiggly_mod) %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred))
```

Adding all predictions to data frame and 'pivoting' so the result is a tidy, 'long' dataset

```{r}
train_df %>% 
  gather_predictions(linear_mod, smooth_mod, wiggly_mod) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_wrap(~model)
```

Quantify the results 

* rmse: room mean square model 
* output seems to follow trends of graphs 
* problem - only done 1 80/20 split, maybe in general line would be ok or maybe wiggly would be ok 
* if get different 80/20 split can you check the results over and over? --> iteration

```{r}
rmse( linear_mod, test_df)
rmse( smooth_mod, test_df)
rmse( wiggly_mod, test_df)
```

## CV iteratively 

use `modelr::crossv_mc`

* training testing splits 100 times 
* want to get rmse all the way through (fitting all 3 models)
* pulling out rows to show which are included in train (just to see what is happening -  not necessary step)
* making train and test tibbles

```{r}
cv_df = 
  crossv_mc(nonlin_df, 100)

cv_df %>% pull(train) %>% .[[1]] %>% as_tibble()
cv_df %>% pull(train) %>% .[[100]] %>% as_tibble()

cv_df = 
  crossv_mc(nonlin_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

Let's fit models

* finding rmse for multiple 80/20 splits
* compare to smooth and wiggly model 
```{r}
cv_df = 
  cv_df %>% 
  mutate(
    linear_mod = map(.x = train, ~lm(y ~ x, data = .x)),
    smooth_mod = map(.x = train, ~gam(y ~ s(x), data = .x)),
    wiggly_mod = map(.x = train, ~gam(y ~ s(x, k = 30), sp = 10e-6, data = .x))
  ) %>% 
  mutate(
    rmse_linear = map2_dbl(.x = linear_mod, .y = test, ~rmse(model = .x, data = .y)),
      rmse_smooth = map2_dbl(.x = smooth_mod, .y = test, ~rmse(model = .x, data = .y)),
      rmse_wiggly = map2_dbl(.x = wiggly_mod, .y = test, ~rmse(model = .x, data = .y))
  )
```

look at output

* tidy df to make plot
* make box plot to see rmse differences

```{r}
cv_df %>% 
  select(.id, starts_with("rmse")) %>% 
  pivot_longer(
    rmse_linear:rmse_wiggly,
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_boxplot()
```

* wanted lowest rmse
* smooth - as expected - has lowest, best predictabilty from rmse 
