---
title: "Linear Models"
output: github_document
---

```{r setup}
library(tidyverse)
library(p8105.datasets)
library(viridis)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE, 
  fig.width = 8, 
  fig.height = 6, 
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "virids"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Load NYC Airbnb data. 

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(stars = review_scores_location / 2) %>% 
  rename(
    borough = neighbourhood_group
  ) %>% 
  filter(borough != "Staten Island") %>% 
  select(price, stars, borough, neighbourhood, room_type)
```

Visualizaitons...
```{r}
nyc_airbnb %>% 
  ggplot(aes(x = stars, y = price)) + geom_point()
```

Fitting linear model ...

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

Look at fit ...

```{r}
summary(fit)
summary(fit)$coef

fit %>% broom::tidy()
```

if you want to present output... broom allows you to present output in a dataframe and then can clean up a dataframe as you normally would 

```{r}
fit %>% 
  broom::tidy() %>% 
  mutate(term = str_replace(term, "borough", "Borough: ")) %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)
```


## Diagnostics
can add residuals from multiple fits - use as a way to compare predicted values/fitted values etc. 

```{r}
modelr::add_residuals(nyc_airbnb, fit) %>% 
  ggplot(aes(x = stars, y = resid)) + geom_point()

modelr::add_residuals(nyc_airbnb, fit) %>% 
  ggplot(aes(x = resid)) + geom_density() +
  xlim(-200, 200)
```

## Interactions? Nesting?

Try to fit a different model ...

```{r}
fit = lm(price ~ stars * borough + room_type * borough, data = nyc_airbnb)

broom::tidy(fit)
```

Try nesting ...

* creating a smaller version of dataset but a little bit easier to work with 
* collapsed everything down into a single data column 

```{r}
nyc_airbnb %>% 
  relocate(borough) %>% 
  nest(data = price:room_type) %>% 
  mutate(
    lm_fits = map(.x = data, ~lm(price ~ stars + room_type, data = .x)), 
    lm_results = (map(lm_fits, broom::tidy))
  ) %>% 
  select(borough, lm_results) %>% 
  unnest(lm_results) %>% 
  filter(term == "stars")
```

Look at neighborhood in Manhattan ...

* density plots - filtering stars: what is the effect of stars keeping room type constant, how much variability is there in the effect of stars 

```{r}
manhattan_lm_results_df = 
nyc_airbnb %>% 
  filter(borough == "Manhattan") %>% 
  select(-borough) %>% 
  relocate(neighbourhood) %>% 
  nest(data = price:room_type) %>% 
  mutate(
    lm_fits = map(.x = data, ~lm(price ~ stars + room_type, data = .x)),
    lm_results = map(lm_fits, broom::tidy)
  ) %>% 
  select(neighbourhood, lm_results) %>% 
  unnest(lm_results)

manhattan_lm_results_df %>% 
  filter(term == "stars") %>% 
  ggplot(aes(x = estimate)) + geom_density()

manhattan_lm_results_df %>% 
  filter(str_detect(term, "room_type")) %>% 
  ggplot(aes(x = neighbourhood, y = estimate)) + geom_point() +
  facet_grid(~term) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


## Logistic regression

```{r}
nyc_airbnb_new = nyc_airbnb %>% 
  mutate(
    expensive_apt = as.numeric(price > 500)
  )
```

Fit logisitc regression for binary outcome.

* glm will assume linear model with continuous outcome
* will run -- fit line through 0 and 1
* need to add "family = binomial()"


by default -- glm will report log odds -- exponentiate to get odds ratio
```{r}
logistic_fit = 
  glm(
    expensive_apt ~ stars + borough,
    data = nyc_airbnb_new, 
    family = binomial())

logistic_fit %>% 
  broom::tidy() %>% 
  mutate(
    term = str_replace(term, "borough", "Borough: "), 
    estimate = exp(estimate)) %>% 
  select(term, OR = estimate, p.value)

nyc_airbnb_new %>% 
  modelr::add_predictions(logistic_fit) %>% 
  mutate(pred = boot::inv.logit(pred))
```

add_predictions --> need to invert logit operation to get probabilities using boot package 
